# Reproducible Research: Peer Assessment 1



##Loading and preprocessing the data
```{r}
unzip(zipfile="repdata_data_activity.zip")
data <- read.csv("activity.csv")
```



###What is mean total number of steps taken per day?
```{r}
library(ggplot2)
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, binwidth=1000, xlab="Total Number of Steps taken per Day")
## Saving Plot1 to file
dev.copy(png, file="plot1.png", height=480, width=480)
dev.off()
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```



###What is the average daily activity pattern?
```{r}
library(ggplot2)
averages <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-Minute Interval") +
    ylab("Average Number of Steps Taken")
## Saving Plot2 to file
dev.copy(png, file="plot2.png", height=480, width=480)
dev.off()
```



####Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
averages[which.max(averages$steps),]
```



##Imputing missing values
####There are many days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.
```{r}
missing <- is.na(data$steps)
# How many missing
table(missing)
```



#### The mean/median for that day, or the mean for that 5-minute interval, etc.
```{r}
# Replace each missing value with the mean value of its 5-minute interval
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (averages[averages$interval==interval, "steps"])
    return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
```



####Let's make a histogram of the total number of steps taken each day and calculate the mean and median total number of steps using the filled data set.
```{r}
total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="Total Number of Steps Taken per Day")
## Saving Plot3 to file
dev.copy(png, file="plot3.png", height=480, width=480)
dev.off()
mean(total.steps)
median(total.steps)
```


####After imputing data the mean and median values are high



###Are there differences in activity patterns between weekdays and weekends?

#
#

####Let's first find the day of the week for each measurement in the dataset. In this part, we use the dataset with the filled-in values.
```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN=weekday.or.weekend)
```


####Making a panel plot containing plots of average number of steps taken on weekdays and weekends
```{r}
averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
    xlab("5-minute interval") + ylab("Number of steps")
## Saving Plot4 to file
dev.copy(png, file="plot4.png", height=480, width=480)
dev.off()
```